# Array of Imports with Monikers

There are variety of import patterns in `box::use()`. The specific pattern: `mod[...]` allows specific qualification of imports, similar to how indexing in R works. Let's first compare the traditional R approach with `{box}`. Here's the following table:

| Aspect | Base R | {box} Approach |
|------------------|--------------------------|----------------------------|
| Loading Method | `library(dplyr)` | `box::use(dplyr)` / `box::use(dplyr[...])` |
| Namespace Impact | Attaches to global namespace | Module-scoped imports |
| Conflict Handling | May cause naming conflicts | Explicit imports prevent conflicts |
| Code Clarity | Dependencies implicit | Dependencies explicitly declared |
| Performance | Loads entire package namespace | More controlled |

Oh, and there's also `require()` function, an "evil-ler" twin of `library()` function, where it won't just attach the *entire* namespace to the search path, it returns boolean (`TRUE` if the package exists; `FALSE` if it doesn't) when executed. There's an inconsistency: it fails silently (“fails silently” means that a function fails without stopping the program or showing a clear error message).

## Importing R Packages and scripts / folders

If the table is not enough to you, this section will further explain. Let's explore the different importing methods:

### Basic Imports

Chapter 1 already shows you the basic ways to import modules using `box::use()`. This function allows multiple imports at once, whether that module is a package, a local module, or packages managed by `{carrier}`. This is not different from the ellipsis (`...`) that allows passing multiple arguments, similar to Python's `*args` and `**kwargs`. Differently, there's a feature in `box::use()`'s bracketing method `mod[]` called "wildcard" in the form of `...`, e.g. `mod[...]`, which will attach [all the names under that namespace]{.underline}. 

Since this allows you to import multiple packages and their functions in a single call, separated by comma, you don't have to bother yourself calling multiple calls, like:

```{r}
#| eval: false
box::use(
    # Packages
    dplyr, ggplot2, stats,
    dplyr[filter, select, mutate], 
    ggplot2[ggplot, geom_point], 
    stats[lm], 
    
    # Local modules
    ./mod, ./folder/mod, 
    ./mod[fn1, ...], 
    ./folder/mod[fn1, ...], 
    
    # Carrier modules
    mod, mod/submodule, 
    mod[fn1, ...], 
    mod/submodule[fn1, ...]
)
```

When importing scripts and modules, you are going to provide the name of the path (should be a literal name, not a string), and/or add prefix `./` that indicates the current path. The use of `../` is allowed as well, but this will be discussed in [Chapter 3.4](3.4-nested-modules.qmd). The rules applied on `{box}`-`{carrier}` modules except you are not allowed to place a prefix since they are packages, not a local module. 

### Importing the entirety

Do not confuse this with `library()`, where it attaches the *entire* namespace to the search path. Throwing all the exports into the search path is often *subjectively* (or *objectively*?) bad and is being discouraged in best software engineering practices. `{box}` has a different: This is the same as Python's `import math` semantics, where instead of loading the package goes to the global environment, the namespace of the package and scripts / folders, including the functions and other objects, such as data frames or constants like `pi` from base R, are encapsulated as *environments*, (another data structure, similar to lists but mutable) with its name, and then the imports are accessed directly with `$` subset operator.

``` r
box::use(
    dplyr,
    ./mod1, 
    ./mod2
)  

dplyr$select(data, col1, col2)
mod1$fn1(mod2$fn3(5))
```

How beautiful this may be? The entire package namespace becomes available as an *environment*. All functions are (and must) be accessed using the package name as a prefix and `$` operator. This is nice as it prevents namespace pollution while maintaining access to all functions. This might not sound useful since you already did that with `pkg::` in base R, this is still useful, especially for local modules and `{box}`-`{carrier}` modules, where you don't want to allow to qualify their names. 

The use of aliases also allowed:

``` r
box::use(
    dpr = dplyr, 
    tdr = tidyr, 
    md1 = ./module1,
    md2 = ./module2
)  

# Usage:
dpr$filter(data, col1 > 0) |> 
    tdr$pivot_longer(cols = col2)

md1$fn1(md2$fn3(5))
```

### Granular Imports with Aliases {spec-imports}

`{box}` absolves namespace clashes, which commonly occur when different packages have functions with the same name.

As you import the names, you are allowed to spice up things a little bit by renaming the original names during the imports. This is superbly useful when you have 2 packages to be used and you want to use them both at once. For instance, the `{dplyr}`'s `filter()` function, and as you attach the `{dplyr}` namespace, it will mask the existing functions from the global namespace, namely the `{stats}`' `filter()` function. This induces namespace clash, and there are times you may not want this happening.

Fortunately with `{box}`, you can load the `filter()` function from the `{dplyr}` namespace through the following:

``` r
box::use(
    dplyr[keep_when = filter],  # dplyr's `filter()` becomes `keep_when()`
    stats[filter_ts = filter]   # stats' filter becomes filter_ts
)

data |> keep_when(col > 0)      
AirPassengers |> filter_ts()      
```

But wait, there's another thing you need to know. When you declare an import under `box::use()`, something like `box::use(pkg = package[fn1])`, two things will happen: this imports `pkg` carrying the namespace `package`, and at the same time will attach `fn1` under `package` namespace. 

::: callout-note
Granular imports in R's `library()` also allowed using `include.only` parameter.

For example:

``` r
library(dplyr, include.only = c("select", "filter"))
```

But still no *alias* gimmicks and doesn't even leverage non-standard evaluation, where it treats `include.only` arguments as an object, called as a name, unlike `box::use()`.

R version 4.4 and above has a shorthand of `library(pkg, include.only = c('fn1', 'fn2'))`: Introducing the `base::use()`.

Example usage:

``` r
use(dplyr, c("select", "filter"))
```

[This is inconsistent, as well, and generally not recommendable.](https://joshuamarie.com/posts/06-load-pkg/#base-use).
:::

### Importing non-syntactic names

The infix operators like `%>%` from `{magrittr}` are considered [non-syntactic names](https://adv-r.hadley.nz/names-values.html#non-syntactic). In R, operators and functions that use special characters (like `+`, `*`, `%>%`, `%in%`, `%*%`, etc.) are called *infix operators* but internally parsed as the usual call in R, e.g. `add(a, b)`. These operators require special handling when importing with `{box}`. By the way, yes, these are still *functions*.

To import functions with special characters, you need to wrap them in backticks:

```{r}
#| echo: true
box::use(
    magrittr[`%>%`, `%T>%`], 
    dplyr[group_by, summarise],
    stats[sd, median]
)

iris %>% 
    group_by(Species) %>% 
    summarise(
        mu = mean(Sepal.Length, na.rm = TRUE), 
        sigma = sd(Sepal.Length, na.rm = TRUE),
        md = median(Sepal.Length, na.rm = TRUE)
    )
```

### Wildcard import

The true equivalent of `library(pkg)` is through `box::use(pkg[...])`. Yet again, `...` are called "ellipsis". The use of `...` sets as a "wildcard", and this, while being *granular*, imports all the namespace within the package (or modules). The Python's equivalent would be `from pkg import *`.

For example:

``` r
box::use(
    dplyr[...]
)
```

## Imports within the function / function call

When we import packages / scripts / folders as modules or import their namespace, did you know the imports are *enclosed* within the scope?

According to the official documentation:

> the effects of `box::use()` are restricted to the current scope: we can load and attach names inside a function, and this will not affect the calling scope (or elsewhere).

Here's an example:

```{r}
#| echo: true
#| code-fold: true
#| collapse: true
#| comment: "#>"
mtcars |> 
    dplyr::reframe(
        {    
            box::use(
                stats[linear_reg = lm, pearson_r = cor],
                purrr[imap_dfc, set_names],
                tibble[tbl = tibble]
            ) 
            
            model = linear_reg(mpg ~ wt)
            coefs = coef(model)
            coef_table = imap_dfc(coefs, \(bi, nm) {
                result = tbl(bi)
                set_names(result, nm)
            })
            
            corr = pearson_r(wt, mpg)
            
            test = summary(model)
            tbl(
                coef_table, 
                corr = corr, 
                rsq = test$r.squared,
                adj_rsq = test$adj.r.squared
            )
        },
        
        .by = cyl
    )
```

::: {.callout-tip title="This code (freely skip this)" collapse="true"}
This code is made to study the type-I error by examining how true the linear relationship between `wt` and `mpg` variables from `mtcars` data, when performing statistical analysis. Here, this code imports within `dplyr::reframe()` function call without making side-effect the current environment. This is great if you create a function with external dependencies available from R packages, or within your scripts / folders.
:::

## Best Practices for Package Imports

But, of course, R packages have strengths, but don't forget their flaws. I will enumerate the *do's* in `{box}` package:

1.  Be Specific with Imports

    Please, do not import everything, unless there's a good use case. The use of *wildcards*, i.e. the "ellipsis" `...` within the granular imports through `[...]` is a shortcut to import the namespace, but you are importing everything here.
    
    ``` r
    box::use(
        dplyr[...]
    )
    ```

    Don't just do this in actual practice, or it will create a mess in the global namespace, just like `library()`. As the *Zen of Python* said: "Explicit is better than implicit."
    
    Instead, import only what you need. Of course, in several times, you only import specific parts of the package only. For instance, when you are aggregating data frame with `{dplyr}`, you often only needs `filter()`, `select()`, `mutate()`, `group_by()`, and `summarise()`. Mind you that there are a total of 293 exported namespaces (will be less than that if you don't count the *pseudo-functions*, such as `across()` and `where()`) within `{dplyr}` package, and for your aggregation task, you only need 5 out of the total exports.

    This approach is better because it is explicit and you can even rename those imports:

    ``` r
    box::use(
        dplyr[filter_df = filter, select, mutate, group_by, summarise]
    )
    ```

    ::: {.callout-tip icon="false" title="Example (Click to expand)" collapse="true"}
    Let's take an example, where you want to calculate the sample size, mean, standard deviation, standard error, and the coefficient of variation across the numeric columns in `iris` dataset:
    
    ```{r}
    #| echo: true
    box::use(
        dplyr[n, mutate, relocate, group_by, summarise], 
        tidyr[melt = pivot_longer, spread = pivot_wider]
    )
    
    iris |> 
        summarise(
            across(
                where(is.numeric), 
                list(
                    mu = \(x) mean(x, na.rm = TRUE), 
                    sigma = \(x) sd(x, na.rm = TRUE)
                ), 
                .names = "{.col}..{.fn}"
            ), 
            n = n()
        ) |> 
        melt(
            cols = everything() & !n,
            names_pattern = "^(.*)\\.\\.(.*)$",
            names_to = c("Variable", "Statistics"),
            values_to = "Est"
        ) |> 
        spread(
            names_from = Statistics, 
            values_from = Est
        ) |> 
        mutate(
            se = sigma / sqrt(n), 
            cv = sigma / mu
        ) |> 
        relocate(
            n, .after =  Variable
        )
    ```
    :::

2.  Group Related Imports

    This is just an opinion. For better clarity, you can explain the imports you are making *within* the function call `box::use()` with comments `#`. This is not unusual, sometimes this is common. Since comments are allowed, you can group the imports by functionality.
    
    ``` r
    box::use(
        # Simple aggregation
        dplyr[filter_df = filter, select, mutate, group_by, summarise],
        tidyr[pivot_longer, pivot_wider],
        
        # To run linear regression and t-test
        stats[linear_reg = lm, welch_ttest = t.test],
        
        # To visualize outputs from linear regression and t-test
        ggplot2[ggplot, geom_point, geom_smooth, geom_box, theme_minimal, aes]
    )
    ```
    
    <!-- ::: callout-note -->
    <!-- The proper documentation is discussed in Chapter 3.2, which it talks about *Package-like Modules*. -->
    <!-- ::: -->

3.  Handle Naming Conflicts

    As discussed [here](#spec-imports), you are allowed to place an alias within the imports, so that the namespace clash will be resolved.
    
    The most prominent example is `dplyr::filter()` and `stats::filter()`.
    
    ``` r
    box::use(
        dplyr[keep_when = filter],
        stats[filter_ts = filter]
    )
    ```

    *Rename them whatever you want to employ clarity.*

### Otherwise

::: {.callout-warning title="Otherwise"}
Trust me, I've been this a long time ago: not following any of those practices. So, if you don't do the following practices, you'll create inconsistent and unpredictable flaws. Keep up the good work and do the best practices.
:::

## Troubleshooting Package Imports

Frankly, I will show you solutions if you have similar errors like the following. The common issues to be found when importing packages happen are:

1.  Obviously, when the R package is not installed. This matter is trivial: The package does not exist in your current environment, and simply just install the packages you need and import them using `box::use()`.

    ``` r
    box::use(pkg[func, ...])
    #> Error in box::use(pkg) : there is no package called ‘pkg’

    install.packages("pkg")
    ```

2.  When the particular imports does not exist in the package namespace or incorrectly name the import:

    ``` r
    box::use(dplyr[nonexistent_function])
    #> Error in box::use(dplyr[nonexistent_function, slct]) : 
         name “nonexistent_function”, “slct” not exported by “dplyr”

    box::use(dplyr[select, filter])
    ```

    `{box}` enforces strict naming. If possible, check out the official documentation of the R package. Check if you are using indeed correct spelled name, and check if the imports does exist in the package namespace

<!-- ## Benefits and history -->

<!-- The `{box}` package is originally a superset of R package import system. Then the author gradually made some breakthrough editions, by interpolating the module system, something that does exists in other languages, but missed out by R throughout the years. And so, you see, using **{box}** for bringing clean module system in R does offers you these advantages: -->

<!-- -   Explicit dependency declaration -->
<!-- -   Better namespace control and reduced conflicts -->

<!-- In the next chapter, we'll explore how to create and reuse your own modules effectively. -->
